---
title: "Final_Project"
author: "Johanne"
date: "2023-01-08"
output: html_document
---

#First load all the necessary libraries.
```{r libraries, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(hrbrthemes)
```

#Read the CSV-file and convert it into a data frame.
```{r}
cause_dk <- read_csv("Disease_DK.csv")

cause_dk
```

#Clean the data frama, but creating three different columns without the semicolons.
#Mutate the data in the new "Count"-column into numeric.
```{r}
cause_disease <- cause_dk %>% 
  separate(1, c("Cause_of_Death", "Year", "Count"))  %>% 
  mutate_at(c("Count"), as.numeric)

cause_disease
```

#Let's make the first graph, showing the causes of deaths in 1990.
```{r}
cause_disease%>% 
  filter(Year %in% c("1990")) %>% 
   group_by(Year,
           Cause_of_Death,
           Count) %>% 
  tally() %>% 
  ggplot(aes(x = Cause_of_Death,
             y = Count,
             fill = Cause_of_Death)) + 
    geom_bar(position="dodge", stat="identity") +
  scale_y_continuous(breaks = seq(0, 100, 10),
                     labels = seq(0, 100, 10)) +
  theme_minimal(base_size = 10) +
    scale_fill_hue(c = 40) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "1990", 
       y = "Count",
       x = "Disease") +
  scale_y_continuous(breaks=seq(0,30000,2500))
```
#	If the line filter(Year %in% c("1990")) is changed into filter(Year %in% c("2019")), then the graph will show the causes of deaths in 2019, making it possible to compare the data from those two years.
```{r}
cause_disease%>% 
  filter(Year %in% c("2019")) %>% 
   group_by(Year,
           Cause_of_Death,
           Count) %>% 
  tally() %>% 
  ggplot(aes(x = Cause_of_Death,
             y = Count,
             fill = Cause_of_Death)) + 
    geom_bar(position="dodge", stat="identity") +
  scale_y_continuous(breaks = seq(0, 100, 10),
                     labels = seq(0, 100, 10)) +
  theme_minimal(base_size = 10) +
    scale_fill_hue(c = 40) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "2019", 
       y = "Count",
       x = "Disease") +
  scale_y_continuous(breaks=seq(0,30000,2500))
```

#Check the development the number of deaths by neoplasms throughout the years.
```{r}
cause_disease%>% 
  filter(Cause_of_Death %in% c("Neoplasms")) %>% 
  filter(Year %in% c("1990", "1992", "1994", "1996", "1998", "2000", "2002", "2004", "2006", "2008", "2010", "2012", "2014", "2016", "2019")) %>% 
  tail(15) %>%
  ggplot( aes(x=Year, y=Count, group=Cause_of_Death)) +
    geom_line( color="#69b3a2") +
    geom_point(shape=21, color="black", fill="#69b3a2", size=3) +
    theme_ipsum() +
  ggtitle("Neoplasms")
```
#What about the development of AIDS?
```{r}
cause_disease%>% 
  filter(Cause_of_Death %in% c("AIDS")) %>% 
  filter(Year %in% c("1990", "1992", "1994", "1996", "1998", "2000", "2002", "2004", "2006", "2008", "2010", "2012", "2014", "2016", "2019")) %>% 
  tail(15) %>%
  ggplot( aes(x=Year, y=Count, group=Cause_of_Death)) +
    geom_line( color="#69b3a2") +
    geom_point(shape=21, color="black", fill="#69b3a2", size=3) +
    theme_ipsum() +
  ggtitle("AIDS")
```
#In the end, a new graph will be made to compare the development throughout the 30 years of the three biggest causes of death in 1990.

```{r}
cause_disease%>% 
  filter(Year %in% c("1990", "2000", "2010", "2019")) %>% 
  filter(Cause_of_Death %in% c("CardiovascularDiseases", "Neoplasms", "ChronicRespiratoryDiseases")) %>%
  group_by(Count,
           Year,
           Cause_of_Death) %>% 
  tally() %>% 
  ggplot(aes(x = Year,
             y = Count,
             fill = Cause_of_Death)) + 
  geom_bar(position="dodge", stat="identity", width=0.8) +
  theme_minimal(base_size = 10) +
  scale_fill_hue(c = 35) +
  scale_y_continuous(breaks=seq(0,30000,2500))
```
